# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r}
data <- read.csv(unz("activity.zip", "activity.csv"),
                 sep=",",
                 header=TRUE,
                 stringsAsFactors=FALSE)
head(data, 10)
```

## What is mean total number of steps taken per day?

### Make a histogram of the total number of steps taken each day.

```{r}
library(ggplot2)
qplot(date, 
      data=data, 
      weight=steps, 
      geom="histogram", 
      xlab="Date", 
      ylab="Total Number of Steps") + 
    geom_histogram(colour="black", fill="red") + 
    theme(axis.text.x=element_text(angle=-90, hjust=0))
```

### Calculate and report the mean and median total number of steps taken per day.

```{r}
mean(na.omit(data$steps))
median(na.omit(data$steps))
```

## What is the average daily activity pattern?

### Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
d <- aggregate(steps ~ interval, data, mean)
plot(d$interval, 
     d$steps, 
     type="l",
     xlab="5-Minute Interval",
     ylab="Average Number of Steps")
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
d[d$steps==max(d$steps),]$interval
```

## Imputing missing values

### Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
sum(is.na(data))
```

### Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r}
cleandata <- function(data){
    aggdata <- aggregate(steps ~ interval, data, mean)
    for(i in data[is.na(data$steps),]$interval){
        data[data$interval == i,]$steps <- aggdata[aggdata$interval == i,]$steps
    }
    data
}
```

### Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
newdataset <- cleandata(data)
head(newdataset)
```

### Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r}
library(ggplot2)
qplot(date, 
      data=newdataset, 
      weight=steps, 
      geom="histogram", 
      xlab="Date", 
      ylab="Total Number of Steps") + 
    geom_histogram(colour="black", fill="red") + 
    theme(axis.text.x=element_text(angle=-90, hjust=0))

mean(newdataset$steps)
median(newdataset$steps)
```

### Do these values differ from the estimates from the first part of the assignment? 

### What is the impact of imputing missing data on the estimates of the total daily number of steps?


## Are there differences in activity patterns between weekdays and weekends?

